---
title: "PS - Oct. Solutions"
author: "JZ"
date: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(haven)
exit <- read_dta("https://github.com/zilinskyjan/R-stata-tutorials/blob/master/homework/31116399_Florida%202018.dta?raw=true")
```

# R Answers

## Part A: Descriptives

1. How many respondents were 65 years old or older? What was their proportion in the sample?

One option:

```{r}
exit %>% count(AGE65)

# Another option:
# exit %>% count(AGE8)
```

Add a column with proportions:

```{r}
exit %>% 
  count(AGE65) %>%
  mutate(proportion = n / sum(n))
```

Another possibility:

```{r}
table(exit$AGE65)
```

```{r}
table(exit$AGE65) / sum(table(exit$AGE65))
```


Even better to run:

```{r}
exit %>% 
  count(AGE65) %>%
  filter(!is.na(AGE65)) %>%
  mutate(proportion = n / sum(n))
```


2. What was the proportion of Hispanic respondents who:

- Lived in cities with pop. over 50,000?
- Lived in suburbs?
- Lived in small cities or rural areas?

(Hint: Look at the variation in the variable labeled `SIZEPLC3`.)

```{r}
exit %>% filter(latino==1) %>%
  group_by(SIZEPLC3) %>% 
  summarize(n = n()) %>%
  mutate(proportion = n / sum(n))
```

Alternative code:

```{r}
exit %>% filter(latino==1) %>%
  group_by(SIZEPLC3) %>% 
  tally() %>%
  mutate(proportion = n / sum(n))
```
```

3. Prepare a table displaying the proportion of voters who said they were first-time (midterm election) voters, broken down by gender.

```{r}
exit %>% group_by(FTVOTER1) %>% tally()
```

```{r}
exit %>% group_by(FTVOTER1,sex) %>% 
    filter(!is.na(FTVOTER1),!is.na(sex)) %>%
    tally() %>%
    mutate(prop = n / sum(n))
```

Why is it important to remove the missing observations from the denominator?

- Most student correctly say that 78 of male voters were first-time voters.
- But we cannot divide 78 by the total number of male respondents (1398).
- Rather, we must divide 78 by 427 (i.e. the number of male respondents for who we know their FT-voting status).

```{r}
exit %>% count(sex)
```

```{r}
exit %>% filter(!is.na(FTVOTER1)) %>% count(sex)
```




4. What was the proportion of Democrats who said in 2018 that Donald Trump should not be impeached and removed from office?

```{r}
exit %>% count(party,IMPEACH1)
```

Let's limit our attention to Democrats:

```{r}
exit %>% count(party,IMPEACH1) %>% filter(party==1)
```

## Part B

1. Is there an association between 2016 vote choice and the type of area where a voter lives? 

```{r}
exit %>% count(SIZEPLC3,VOTE2016)
```

Run a regression:

```{r}
exit$votedForCliton2016 <- ifelse(exit$VOTE2016==1,1,0)

model_vote <- lm(votedForCliton2016 ~ factor(SIZEPLC3), data= exit)
```


2. What percentage of voters said that when choosing their candidate for the US Senate, "Donald Trump was not a factor"?

```{r}
exit %>% count(fortrump)
```


3. What percentage voters said that, in their vote for the US Senate, Bill Nelson's vote against Brett Kavanaugh's confirmation "not a factor at all?"

```{r}
exit %>% count(KAVFL18)
```

## Part C:

```{r}
exit$motivated <- ifelse(exit$RUSSIA18==2,1,0)
```

```{r}
exit <- exit %>% mutate(politcally_motivated = ifelse(RUSSIA18==2,1,0))
```


Note: Republicans are `party==2`:

```{r}
mod <- lm(motivated ~ factor(party) + factor(AGE8),
              data = exit %>% filter(!is.na(RUSSIA18)))
```

```{r}
summary(mod)
```

What are the key things to notice here?

* Higher values of `party` do not mean that a respondent is "more Republican".
* The values are not ordered in a meaningful way (1 = Democrat; 2 = Republican; 3 = "Independent).
* Even if the values were ordered in an ideological manner, it still doesn't mean that treating the variable as continuous is necessarily a defensible choice.
* The values are unordered, so you *must* include a series of indicator/dummy variables, if you wish to control for partisanship.
* One way to achieve that is to add `factor(party)` to your regression forumula.



\pagebreak  

# STATA Answers

`tab AGE8 or tab AGE65`

`gen hispanic = (latino==1) if !mi(latino)`

`tabstat hispanic, by(SIZEPLC3)`

`tab FTVOTER1 sex`

`tab party IMPEACH1, row nof`



## Part B:

B1: `tab SIZEPLC3 VOTE2016 [aw=weight], row nof  `

B2: `tab fortrump`

B3: `tab KAVFL18`

## Part C:

`gen motivated = (RUSSIA18==2) if !mi(RUSSIA18)`

`reg motivated i.AGE8 ib3.party`

It is important to note that:

* Higher values of party do not mean that a respondent is "more Republican".
* The values are of the variable not ordered in a meaningful way (1 = Democrat; 2 = Republican; 3 = "Independent). And even if the values were ordered ideologically, it still doesn't mean that treating the variable as continuous is necessarily a defensible choice.
* Given that the values are unordered, then you *must* include a series of indicator/dummy variables.
* One way to achieve that is to use the "i." operator.

